---
title: "Volunteers in NESTA experiment \\linebreak  Technical Report"
author: "Rafal Urbaniak"
output:
  pdf_document:
    keep_tex: yes
  number_sections: yes
  df_print: kable
includes:
  in_header: Rafal_latex8.sty
html_document:
  df_print: paged
word_document: default
classoption: dvipsnames, enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
---



```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')

#libraries used
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(viridis)
library(rethinking)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(GGally)
library(dagitty)
library(reshape)
library(lubridate)
library(formatR)
library(ggrepel)


```
  
  
The winning model, given our model selection method, is specified as follows:

\begin{align*}
\mathsf{interventions} \sim \mathsf{NegativeBinomial} (\lambda,\phi) \\
log(\lambda) = l_\mathsf{volunteerID[i]} + enth_\mathsf{ volunteerID[i]} \times \mathsf{daysOfProject} + comp_\mathsf{volunteerID[i]} \times \mathsf{competition}\\  
l_\mathsf{ volunteerID[i] }  \sim \mathsf{Norm}(lbar,lsigmabar) \\
lbar \sim \mathsf{Norm}(2, .9)\\
lsigmabar, enthsigmabar, compsigmabar \sim  \mathsf{Exp}(.5) \\
enth _\mathsf{ volunteerID[i] }  \sim \mathsf{Norm}(enthbar, enthsigmabar)\\
comp_\mathsf{ volunteerID[i] } \sim \mathsf{Norm}(compbar, compsigmabar) \\
enthbar, compbar \sim  \mathsf{Norm}(0, .3)\\
 \phi =  puser_\mathsf{ volunteerID[i] }  \\
 puser_\mathsf{ volunteerID[i] } \sim \mathsf{Exp}(1)
\end{align*}


Intuitively, volunteer interventions are assumed to have negative binomial distribution around their own expected value $\lambda$ and individualized dispersion parameters $\phi$. On each day each a user has their own daily expected value, which is determined by the following factors:
\begin{itemize}
\item First, there's user's individual baseline activity for the whole treatment period, $l_\mathsf{ volunteerID[i] }$.
\item next, each user has their own dispersion parameter,  $puser_\mathsf{ volunteerID[i] }$.
\item then, there is (usually dwindling) enthusiasm: the impact of time on that user, $enth_\mathsf{volunteerID[i]} $ to be (after exponentiation) multiplied by the number of days that have passed since the experiment started,
\item finally, we have the impact that the presence of competitions made on a user, $comp_\mathsf{ volunteerID[i] }$, which (after exponentiation) becomes the activity multiplier to be applied during competitions only.
\end{itemize}

\noindent Moreover, the model is hierarchical: the individual level parameters are drawn from distributions whose parameters are in turn to be estimated as well. Thus, $lbar$ is the overall baseline for the whole group, $enthbar$ is the overall estimated group enthusiasm coefficient, and $compbar$ is the overall estimated competition impact coefficient (all of them come with their own nuisance sigma parameters).



\noindent All of these parameters are given priors in a manner analogous to the introduction of priors for the other time series models, as explained in the appendix.^[Interestingly, if we are interested in the causal effect of competitions, we should not use an auto-regressive predictor. If we auto-regress on a lag in the $[1,7]$ range, for some days we will be conditioning on interventions conducted during the same competition, which will already contain some information about the impact of that competition. In other words, auto-regression with short lags would lead to post-treatment bias. On the other hand, auto-regression with longer lags would either lead to dropping a lot of data in the beginning (where lagged information is not available), or degenerate the analysis by using 0s for missing lagged values in a long initial period. All this without much gain, as we have already inspected null models with  auto-regression with large lags and they do not lead to performance improvement.] 

Raw data and daily means are illustrated in Figure \ref{fig:volunteersBasic}, and the individualized totals with the key coefficients based on the trained model are illustrated in Figure \ref{fig:volunteersModel}.




```{r volunteersVis,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE, tidy = FALSE}
volunteersLong <- read_csv(file = "nestaVolunteers/volunteersLong.csv")
competitionModel<- readRDS( "nestaVolunteers/competitionModel.rds")
volunteersJoint <- readRDS( file = "nestaVolunteers/datasets/volunteersJoint.rds")

profile <- c("low", "low", "low", "low",
             "initial", "initial", "initial",  "selective",
             "dynamic", "initial","selective","selective",
             "dynamic","dynamic", "selective")

volunteers <- unique(volunteersJoint$volunteer)
volunteersIDs <- unique(volunteersJoint$volunteerID)

volunteersSummary <- volunteersJoint %>%
  group_by(volunteer) %>%
  summarise(totalActivity = sum(interventions))

volunteersSummary$volunteerID <- as.factor(volunteersIDs)

volunteersSummary$profile <- factor(profile, levels = c("dynamic","active", "selective", "initial", "low"))
volunteersSummary <- as.data.frame(volunteersSummary)

meanRunning <- volunteersJoint  %>%
  group_by(Date) %>%
  summarise(competition = mean(competition), meanInterventions = mean(interventions))

volunteersDailyMeansPlot <- ggplot(data = meanRunning)+ 
  geom_violin(aes(x= as.factor(competition), y = meanInterventions,
                  color = as.factor(competition)))+
  theme_tufte(base_size = 10)+xlab("competition")+
  ylab("mean daily interventions")+
  ggtitle("Daily intervention means were higher during competitions")+
  geom_jitter(width = .3, aes(x= as.factor(competition),
              y = meanInterventions, 
              color =  as.factor(competition), shape = Date < "2020-07-16",
              size = meanInterventions)
              , alpha = .6)+
  theme(plot.title.position = "plot",legend.position = c(0.09, 0.8), legend.key.size = unit(.3, "cm") )+
  scale_size(range = c(1,4))+guides(size = "none", color = "none")+
  scale_color_manual(values = c("orangered", "grey"),
                     labels = c("yes", "no"), name = "competition",
                     breaks = c("1", "0"))+
  guides(shape = guide_legend(title = "First week"))




volunteersDailyPlot <- ggplot(volunteersJoint)+geom_point(aes(x = Date, y = interventions, color = as.factor(competition), size = interventions), alpha = .8)+
  theme_tufte(base_size = 10)+
  scale_color_manual(values = c("orangered", "grey"), labels = c("yes", "no"), name = "competition", breaks = c("1", "0"))+xlab("")+
  theme(plot.title.position = "plot",legend.position = c(0.09, 0.9), legend.key.size = unit(.3, "cm") )+ 
  scale_size(range = c(.5,2))+guides(size = "none")+labs(title = "After initial enthusiasm, volunteer engagement was higher during competitions")+ 
  geom_line(data = meanRunning, aes(x = Date, y = meanInterventions))

```





\begin{figure}
```{r fig:volunteersBasic5,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
grid.arrange(volunteersDailyPlot, volunteersDailyMeansPlot, ncol = 1)
```
\caption{Daily individual voilunteer intervention counts accross time with competition periods marked (top) and daily group intervention means grouped by whether a competition was ongoing (bottom). Note most of high means in the non-competition period are in the first week.}
\label{fig:volunteersBasic}
\end{figure}

```{r fig:volunteersModelPrep,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE, tidy = FALSE}

annotsize <- 2.5

sumsPlot  <- ggplot(volunteersSummary) +geom_bar(aes(x = reorder(as.factor(volunteerID), totalActivity)
                                        , y = totalActivity, fill = profile), stat = "identity")+
  coord_flip()+theme_tufte(base_size = 8)+ylab("total interventions")+xlab("")+
  ggtitle("Total engagement")+
  scale_fill_manual(values = c("darkred", "orangered", "darkgoldenrod", "bisque2", "azure4"),
                    labels = c("active, more during competitions","active throughout" , "active only during competitions", "initial enthusiasm died", "almost no activity"))+
  theme(plot.title.position = "plot", legend.key.size = unit(.25, "cm"),legend.key.width = unit(.3,"cm") , legend.position = c(.35, 1.15),
        legend.title = element_blank())+xlab("volunteer ID")


precis <- precis(competitionModel, depth = 2)
lIndividual <- precis[grep("l\\[",rownames(precis)),]
lGroup <- exp(precis[grep("lbar",rownames(precis)),])

lmean <- exp(lIndividual$mean)
llow <- exp(lIndividual$`5.5%`)
lhigh <- exp(lIndividual$`94.5%`)

eIndividual <- precis[grep("enth\\[",rownames(precis)),]
enthmean <- exp(eIndividual$mean)
enthlow <- exp(eIndividual$`5.5%`)
enthhigh <- exp(eIndividual$`94.5%`)

eGroup <- exp(precis[grep("enthbar",rownames(precis)),])

cIndividual <- precis[grep("comp\\[",rownames(precis)),]
cmean <- exp(cIndividual$mean)
clow <- exp(cIndividual$`5.5%`)
chigh <- exp(cIndividual$`94.5%`)

cGroup <- exp(precis[grep("compbar",rownames(precis)),])

volunteersDF <- data.frame(volunteer = volunteers, totalActivity = 
                             volunteersSummary$totalActivity ,
          profile = volunteersSummary$profile , 
          profileID = as.integer(volunteersSummary$profile ),
           volunteerID = as.integer(as.factor(volunteers)),
           lmean = lmean, llow = llow, lhigh = lhigh,
           emean = enthmean, elow = enthlow, ehigh = enthhigh,
           cmean = cmean, clow = clow, chigh = chigh)

lambdaPlot <- ggplot(volunteersDF)+geom_pointrange(aes(x = reorder(as.factor(volunteerID), totalActivity),y = lmean, ymin  = llow, ymax = lhigh,
                                                       color = profile))+
coord_flip()+theme_tufte(base_size = 8)+xlab("")+ylab(expression(lambda))+theme(axis.text.y =
                element_blank(), axis.ticks.y = element_blank())+ggtitle("Daily baseline")+
  scale_color_manual(values = c("darkred", "orangered", "darkgoldenrod", "bisque2", "azure4"),
                    labels = c("active, more during competitions","active throughout" , 
                               "active only during competitions", "initial enthusiasm died", "almost no activity"))+
  guides(color = "none")+geom_hline(yintercept = lGroup$mean, alpha = .8)+
  annotate("rect", ymin = lGroup$`5.5%`, ymax = lGroup$`94.5%`,
           xmin = 0, xmax = 16.5,
           alpha = .2,fill = "grey")+
  annotate("label", label =  round(lGroup$mean,2), y = lGroup$mean+1, x = 16, size = annotsize)

enthusiasmPlot <- ggplot(volunteersDF)+geom_pointrange(aes(x = reorder(as.factor(volunteerID), totalActivity),y = emean, ymin  = elow, ymax = ehigh,
                                                       color = profile))+
  coord_flip()+theme_tufte(base_size = 8)+xlab("")+ylab("multiplier per day passed")+theme(axis.text.y =
                                                                       element_blank(), axis.ticks.y = element_blank())+
  ggtitle("Enthusiasm change per day")+
  scale_color_manual(values = c("darkred", "orangered", "darkgoldenrod", "bisque2", "azure4"),
                     labels = c("active, more during competitions","active throughout" , 
                                "active only during competitions", "initial enthusiasm died", "almost no activity"))+
  guides(color = "none")+geom_hline(yintercept = 1, alpha = .6, color = "grey", lty = 2)+
  geom_hline(yintercept = eGroup$mean, alpha = .9)+
  annotate("rect", ymin = eGroup$`5.5%`, ymax = eGroup$`94.5%`,
           xmin = 0, xmax = 16.5,
           alpha = .2,fill = "grey")+
  annotate("label", label =  round(eGroup$mean,2), y = eGroup$mean+.02, x = 16, size = annotsize)

competitionPlot <- ggplot(volunteersDF)+geom_pointrange(aes(x = reorder(as.factor(volunteerID), totalActivity),y = cmean, ymin  = clow, ymax = chigh,
                                                           color = profile))+
  coord_flip()+theme_tufte(base_size = 8)+xlab("")+ylab("multiplier if competition is on")+theme(axis.text.y =
                                                                                element_blank(), axis.ticks.y = element_blank())+
  ggtitle("Impact of competitions")+
  scale_color_manual(values = c("darkred", "orangered", "darkgoldenrod", "bisque2", "azure4"),
                     labels = c("active, more during competitions","active throughout" , 
                                "active only during competitions", "initial enthusiasm died", "almost no activity"))+
  guides(color = "none")+geom_hline(yintercept = 1, alpha = .6, color = "grey", lty = 2)+
  geom_hline(yintercept = cGroup$mean, alpha = .9)+
  annotate("rect", ymin = cGroup$`5.5%`, ymax = cGroup$`94.5%`,
           xmin = 0, xmax = 16.5,
           alpha = .2,fill = "grey")+
  annotate("label", label =  round(cGroup$mean,2), 
           y = cGroup$mean+1, x = 16, size = annotsize)

sumsPlotGrob <- ggplotGrob(sumsPlot)
lambdaPlotGrob <- ggplotGrob(lambdaPlot)
enthusiasmPlotGrob <- ggplotGrob(enthusiasmPlot)
competitionPlotGrob <- ggplotGrob(competitionPlot)



```




\begin{figure}
```{r fig:volunteersModel17,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "250%", out.extra='angle=90', warning = FALSE, message = FALSE}
ggplot(data.frame(a=1)) + xlim(1, 38) + ylim(1, 10)+theme_void()+
annotation_custom(sumsPlotGrob, xmin = 1, xmax = 10, ymin =1.2, ymax = 9.56)+
annotation_custom(lambdaPlotGrob, xmin = 10, xmax = 20, ymin =1, ymax = 10)+
annotation_custom(enthusiasmPlotGrob, xmin = 18, xmax = 28, ymin = 1, ymax = 10)+
annotation_custom(competitionPlotGrob, xmin = 28.5, xmax = 38.5, ymin =1, ymax = 10)
```
\caption{Volunteer total engagement with their daily baseline and multipliers for enthusiasm and impact of competition. Pointranges represent individual level coefficients, group coefficients are represented by black lines with shaded 89\% HPDI areas.}
\label{fig:volunteersModel}
\end{figure}





